\textbf{Computational Fuzzy Extractors}

### 2013-06-18 trunk/CompFuzzExt.tex
edited decode and its efficiency analysis

\subsection{Efficiency of \consref{cons:informal construction}}
\label{sec:time main construction}
%\subsection{Analysis of Construction~\ref{cons:LWESecureSketch}}
\consref{cons:informal construction} is only useful if $\decode_t$ can be efficiently implemented.  We need a decoding algorithm for a random linear code with $t$ errors that runs in polynomial time.  We present a simple $\decode_t$.  An improved decoding algorithm can replace our algorithm, which will increase our correcting capability and improve \consref{cons:informal construction}.

\begin{construction}
\label{cons:decoding algorithm} We consider a setting of $(n, m, q, \chi)$ where $m\geq 3n$.  We describe $\decode_t$:
\begin{enumerate}
\item Input $\vA , \vB = \vA X + W - W'$
\item Randomly select rows without replacement $i_1,..., i_{2n}\leftarrow [1,m]$.  
\item Restrict $\vA, \vB$ to rows $i_1,...,i_{2n}$; denote these $\vA_{i_1,...,i_{2n}}, \vB_{i_1,...,i_{2n}}$.
\item Find $n$ rows of $\vA_{i_1,..., i_{2n}}$ that are linearly independent.  
If no such rows exist go to step (2).
\item Denote by $\vA', \vB'$ the restriction of $\vA_{i_1,..., i_{2n}}, \vB_{i_1,..., i_{2n}}$ (respectively) to these rows. Compute $X' = (\vA')^{-1}B'$.  
\item If $\vB- \vA X'$ has more than $t$ nonzero coordinates, go to step (2).
\item Output $X'$.
\end{enumerate}
\end{construction}

Each step is computable in time $O(n^3)$. 
For $\decode_t$ to be efficient, we need $t$ to be small enough so that  with probability at least $\frac{1}{\poly(n)}$, none of the $2n$ rows  selected  in step 2 have errors (i.e., so that $W$ and $W'$ agree on those rows), and $\vA_{i_1,...,i_{2n}}$ has rank  $n$.  If this happens, then $X'=X$, and the algorithm terminates.  However, we also need to ensure correctness: we need to make sure that if $X'\neq X$, we detect it in step 6.  This detection will happen if $B-\vA X' = \vA (X-X')+(W-W')$ has more than $t$ nonzero coordinates.  It suffices to ensure that $\vA (X-X')$ has at least $2t+1$ nonzero coordinates (because at most $t$ of those can be zeroed out by $W-W'$), which happens whenever the code generated by $\vA$ has distance $2t+1$.

Setting $t = O(\frac{m}{n}\log n)$ is sufficient to ensure efficiency.    Random linear codes have distance at least $O(\frac{m}{n}\log n)$ with probability $1-e^{-\Omega(n)}$ (the exact statement is in \corref{cor:code high distance}), so this also ensures correctness.
The formal statement is below~(proof in \secref{sec:proof lem i t poly time}):
%  We provide the results in \lemref{lem:i t constant time} and \lemref{lem:i t poly time} respectively~(proofs in Sections~\ref{sec:proof lem i t constant time},~\ref{sec:proof lem i t poly time}).  \lnote{can we combine these lemmas into one?  It's hard for the reader to compare line-by-line to see where the differences are.  More errors seems better; we could simply point out that constant number of erros leads to better running time, right in the statement of the lemma or right after it}
%\begin{lemma}[Efficiency of $\decode_t$ when $t\leq m/n-1$]
%\label{lem:i t constant time}
%Let $\vA, \vA X+E$ be a  $(n,m, q, \chi)$-LWE instance that is a unique witness relation~(except with negligible probability).  Assume $d(E, E')\leq t$ where $t \leq (m-n)/n$.  Then $\decode_t$ runs in expected time $O(n^3\log q+ s_{ver})$ on input distribution $\vA, \vA X+E - E'$ and outputs $X' =X$ for all but a negligible fraction of inputs over $\vA, \vA X+E - E'$.
%\end{lemma}
\begin{lemma}[Efficiency of $\decode_t$ when $t\leq d (m/n-2)\log n$]
\label{lem:i t poly time}
%Let $\vA, \vA X+W$ be a  $\distLWE_{n,m, q, \chi}$ instance where $m\geq 3n$.  Assume that $W$ is split into $m$ blocks of length $b$ and let $\dis$ be the Hamming distance over alphabet $2^b$. 
Let $d$ be a positive constant and assume that $\dis(W, W')\leq t$ where $t\leq d(\frac{m}{n}-2)\log n$.  Then $\decode_t$ runs in expected time $O(n^{4d+3})$ operations in $\Fq$ (the expectation is taken over the input distribution\bnote{this is probably wrong, because we have a nonzero probability of an infinite running time when $A$ has rank $<n$} $\vA, \vA X+W - W'$).  It outputs $X$ with probability $1-e^{-\Omega(n)}$ (this probability is over the choice of the random matrix  $\vA$).
\end{lemma}
\textbf{Note:}
Even though we are correcting $O(\log n)$ errors, this corresponds to a super polynomial number of possible error patterns; the running time of $\decode_t$ is significantly better than exhaustive search over these error patterns. \lnote{Should we clarify this --- this is about constant errors rather than logarithmic: Furthermore, if we consider the setting of $t\leq m/n-2$, then $\decode_t$ will run in expected time $O(n^3)$ field operations (this does not immediately follow from \lemref{lem:i t poly time}, but requires a simple reanalysis).  Our main theorems get  better parameters when  $t\leq m/n-2$.  }

@@

\begin{figure}[p]
\begin{framed}
\begin{enumerate}
\item Input $\vA', \vb'$, where $\vA' \overset{\$} \leftarrow \Fq^{m\times n}$ and $\vb'$ is either uniform over $\Fq^m$ or $\vb' = \vA'\vx' +\ve'$.
\item Randomly generate $\vect{R} \overset{\$}\leftarrow \Fq^{\alpha \times n}$. Initialize $\vQ \in \Fq^{n\times (\alpha+\beta)}$ to be the zero matrix.
\item Let $\vb^* = (\vb', b^*_{m+1}, \ldots,b^*_{m+\alpha})$, where $(b^*_{m+1}, \ldots, b^*_{m+\alpha} )\overset{\$} \leftarrow \Fq^\alpha$.\label{step:b generation}
\item Randomly generate $\vect{S} \overset{\$}\leftarrow \Fq^{\alpha \times (\alpha+\beta)}$.
		\subitem If $\rank(\vect{S})<\alpha$, stop and output a random bit.
\item Find a set of $\alpha$ linearly independent columns in $\vS$.  Let $T$ be the set of these indices.
\item For all $x_i$ where $n\leq i$ and $i\not \in T$:\label{step:fill in matrix}
\subitem Generate $x_i\overset{\$}\leftarrow \Fq$.  
\subitem For $j=1,..., m$:
\subsubitem Generate $\vQ_{j, i}\overset{\$}\leftarrow \Fq$
\subsubitem Set $b_j^* = b_j^* + \vQ_{j, i} x_i$.
\item Initialize $\vA^*  = \left(\begin{array}{c | c}\vA' & \vQ\\\hline \vR & \vS\end{array}\right)$.
\item \label{step:randomization}
For {$i=1,..., m$}:
\subitem Randomly generate $\gamma_i \leftarrow \Fq^{1 \times \alpha}$.
\subitem Set $\va_{i} \leftarrow \va^*_{i}+\gamma_i (\vR||\vS)$
\subitem Set $b_i \leftarrow b^*_i + \gamma_i (b^*_{m+1},..., b^*_{m+k})$
\item Run $D$ on input $\vA, \vb$.
\item Output $D$'s output.
\end{enumerate}
\end{framed}
\caption{A PPT $D'$ that distinguishes LWE using distinguisher for LWE w/ block fixing source}
\label{fig:perfectLWEreduction}
\end{figure}
The distinguisher $D'$ has an advantage when $S$ is of rank $\alpha$.  This occurs with overwhelming probability~(techniques from Cooper~\cite{cooper2000rank}):
\begin{claim}
\label{cl:full rank matrix}
Let $\vS \overset{\$}\leftarrow \Fq^{\alpha \times (\alpha+\beta)}$ be randomly generated.  Then $\Pr[\rank(\vS)=\alpha]\geq  1- \ngl(n)$.
\end{claim}
\begin{proof}
The probability that all $\alpha$ rows are linearly independent is the successive probability that each row is linearly independent of the previous rows, for the $i$th row this probability is $q^{i-1}/q^{\alpha+\beta}$.  By union bound and $\beta = \omega(1)$, the probability that $\vS$ is not full rank is:
\begin{align*}
\Pr[\rank(\vS) < \alpha] &= \sum_{i=1}^{\alpha} \frac{q^{i-1}}{q^{\alpha+\beta}}
\leq \alpha \frac{q^{\alpha}}{q^{\alpha+\beta} }=\ngl(n)\\
\end{align*}
Thus, $\Pr[\rank(\vS) = \alpha] \geq 1-\ngl(n)$.
\end{proof}
\begin{corollary}
$
|\Pr[D(\vA, \vA\vx+\ve) = 1  | \rank(\vS) = \alpha] - \Pr[D(\vA, \vu) =1 | \rank(\vS) = \alpha]|> \epsilon - \ngl(n).
$
\end{corollary}

@@

\subsection{Proof of \lemref{lem:i t poly time}}
\label{sec:proof lem i t poly time}
\begin{proof}
Let $d$ be a positive constant and $t = d\left( \frac{m}{n}-2\right)\log n$.  
By \corref{cor:code high distance}, the code generated by $\vA$ has distance at least $2t+1$ with probability $1-e^{-\Omega(n)}$.  Thus, if $X\neq X'$, step (6) will detect it with probability at least $1-e^{-\Omega(n)}$.  This probability is over the coins of $\gen$ and thus does not grow with the number of iterations in $\decode_t$.  We consider the expected number of times step 2 is executed when $t\leq d\left(\frac{m}{n}-2\right)\log n$. 
An iteration of $\decode_t$ succeeds when all selected rows of $\vA_{i_1,..., i_{2n}}, B_{i_1,..., i_{2n}}$ have no error and when  $\vA_{i_1,..., i_{2n}}$ has rank $n$.  The event that $\vA_{i_1,..., i_{2n}}$ has rank $n$ is independent of event that the rows $i_1,...,i_{2n}$ have no errors.  The probability of each selected row having an error is at most $\frac{t}{m - i}$ where $i$ is the number of rows already selected.  That is,
\begin{align*}
\Pr[i_1,..., i_{2n}\text{ have no errors}]&\geq \prod_{i=0}^{2n-1}\left(1 - \frac{t}{m-i}\right)\geq \prod_{i=0}^{2n-1}\left( 1-\frac{d\left(\frac{m}{n}-2\right)\log n}{m-i}\right)\\
&\geq  \prod_{i=0}^{2n-1}\left( 1-\frac{d\log n}{n}\left(\frac{m-2n}{m-i}\right)\right)\geq \prod_{i=0}^{2n-1}\left( 1-\frac{d\log n}{n}\right) \\
&= \left(1-\frac{d\log n}{n}\right)^{2n}  = \left(\left(1-\frac{d\log n}{n}\right)^{\frac{n}{d\log n}}\right)^{2d\log n}\geq \frac{1}{4^{2d\log n}} = \frac{1}{n^{4d}}
\end{align*}
Now, we consider the probability  $\vA_{i_1,..., i_{2n}}$  has rank at least $n$.  This probability is at least $1-e^{-\Omega(n)}$~(see \clref{cl:full rank matrix}).  Thus, the probability that a given iteration succeeds $\frac{1}{n^{4d}} - e^{-\Omega(n)}$.  
Thus, the expected number of iterations in step 2 at most  $2n^{4d}$.  We assume all matrix operations can be completed in  $O(n^3)$ operations over $\Fq$, giving a total of expected $O(n^{4d+3})$ operations over $\Fq$.  
\end{proof}

>>>>

\subsection{Efficiency of \consref{cons:informal construction}}
\label{sec:time main construction}
%\subsection{Analysis of Construction~\ref{cons:LWESecureSketch}}
\consref{cons:informal construction} is useful only if $\decode_t$ can be efficiently implemented.  We need a decoding algorithm for a random linear code with $t$ errors that runs in polynomial time.  We present a simple $\decode_t$ that runs in polynomial time and can correct
correcting $O(\log n)$ errors (note that this corresponds to a superpolynomial number of possible error patterns).
This algorithm is only a proof of concept, and neither the algorithm nor its analysis have been optimized for constants. An improved decoding algorithm can replace our algorithm, which will increase our correcting capability and improve \consref{cons:informal construction}.

\begin{construction}
\label{cons:decoding algorithm} We consider a setting of $(n, m, q, \chi)$ where $m\geq 3n$.  We describe $\decode_t$:
\begin{enumerate}
\item Input $\vA , \vB = \vA X + W - W'$
\item Randomly select rows without replacement $i_1,..., i_{2n}\leftarrow [1,m]$.  
\item Restrict $\vA, \vB$ to rows $i_1,...,i_{2n}$; denote these $\vA_{i_1,...,i_{2n}}, \vB_{i_1,...,i_{2n}}$.
\item Find $n$ rows of $\vA_{i_1,..., i_{2n}}$ that are linearly independent.  
If no such rows exist, output $\perp$ and stop.
\item Denote by $\vA', \vB'$ the restriction of $\vA_{i_1,..., i_{2n}}, \vB_{i_1,..., i_{2n}}$ (respectively) to these rows. Compute $X' = (\vA')^{-1}B'$.  
\item If $\vB- \vA X'$ has more than $t$ nonzero coordinates, go to step (2).
\item Output $X'$.
\end{enumerate}
\end{construction}

Each step is computable in time $O(n^3)$. 
For $\decode_t$ to be efficient, we need $t$ to be small enough so that  with probability at least $\frac{1}{\poly(n)}$, none of the $2n$ rows  selected  in step 2 have errors (i.e., so that $W$ and $W'$ agree on those rows).  If this happens, and $\vA_{i_1,...,i_{2n}}$ has rank  $n$ (which is highly likely), then $X'=X$, and the algorithm terminates.  However, we also need to ensure correctness: we need to make sure that if $X'\neq X$, we detect it in step 6.  This detection will happen if $B-\vA X' = \vA (X-X')+(W-W')$ has more than $t$ nonzero coordinates.  It suffices to ensure that $\vA (X-X')$ has at least $2t+1$ nonzero coordinates (because at most $t$ of those can be zeroed out by $W-W'$), which happens whenever the code generated by $\vA$ has distance $2t+1$.

Setting $t = O(\frac{m}{n}\log n)$ is sufficient to ensure efficiency.    Random linear codes have distance at least $O(\frac{m}{n}\log n)$ with probability $1-e^{-\Omega(n)}$ (the exact statement is in \corref{cor:code high distance}), so this also ensures correctness.
The formal statement is below~(proof in \secref{sec:proof lem i t poly time}):
%  We provide the results in \lemref{lem:i t constant time} and \lemref{lem:i t poly time} respectively~(proofs in Sections~\ref{sec:proof lem i t constant time},~\ref{sec:proof lem i t poly time}).  \lnote{can we combine these lemmas into one?  It's hard for the reader to compare line-by-line to see where the differences are.  More errors seems better; we could simply point out that constant number of erros leads to better running time, right in the statement of the lemma or right after it}
%\begin{lemma}[Efficiency of $\decode_t$ when $t\leq m/n-1$]
%\label{lem:i t constant time}
%Let $\vA, \vA X+E$ be a  $(n,m, q, \chi)$-LWE instance that is a unique witness relation~(except with negligible probability).  Assume $d(E, E')\leq t$ where $t \leq (m-n)/n$.  Then $\decode_t$ runs in expected time $O(n^3\log q+ s_{ver})$ on input distribution $\vA, \vA X+E - E'$ and outputs $X' =X$ for all but a negligible fraction of inputs over $\vA, \vA X+E - E'$.
%\end{lemma}
\begin{lemma}[Efficiency of $\decode_t$ when $t\leq d (m/n-2)\log n$]
\label{lem:i t poly time}
%Let $\vA, \vA X+W$ be a  $\distLWE_{n,m, q, \chi}$ instance where $m\geq 3n$.  Assume that $W$ is split into $m$ blocks of length $b$ and let $\dis$ be the Hamming distance over alphabet $2^b$. 
Let $d$ be a positive constant and assume that $\dis(W, W')\leq t$ where $t\leq d(\frac{m}{n}-2)\log n$.  Then $\decode_t$ runs in expected time $O(n^{4d+3})$ operations in $\Fq$.  It outputs $X$ with probability $1-e^{-\Omega(n)}$ (this probability is over the choice of the random matrix  $\vA$ and random choices made by $\decode_t$).
\end{lemma}

@@

\begin{figure}[p]
\begin{framed}
\begin{enumerate}
\item Input $\vA', \vb'$, where $\vA' \overset{\$} \leftarrow \Fq^{m\times n}$ and $\vb'$ is either uniform over $\Fq^m$ or $\vb' = \vA'\vx' +\ve'$.
\item Randomly generate $\vect{R} \overset{\$}\leftarrow \Fq^{\alpha \times n}$. Initialize $\vQ \in \Fq^{n\times (\alpha+\beta)}$ to be the zero matrix.
\item Let $\vb^* = (\vb', b^*_{m+1}, \ldots,b^*_{m+\alpha})$, where $(b^*_{m+1}, \ldots, b^*_{m+\alpha} )\overset{\$} \leftarrow \Fq^\alpha$.\label{step:b generation}
\item Randomly generate $\vect{S} \overset{\$}\leftarrow \Fq^{\alpha \times (\alpha+\beta)}$.
		\subitem If $\rank(\vect{S})<\alpha$, stop and output a random bit.
\item Find a set of $\alpha$ linearly independent columns in $\vS$.  Let $T$ be the set of these indices.
\item For all $x_i$ where $n\leq i$ and $i\not \in T$:\label{step:fill in matrix}
\subitem Generate $x_i\overset{\$}\leftarrow \Fq$.  
\subitem For $j=1,..., m$:
\subsubitem Generate $\vQ_{j, i}\overset{\$}\leftarrow \Fq$
\subsubitem Set $b_j^* = b_j^* + \vQ_{j, i} x_i$.
\item Initialize $\vA^*  = \left(\begin{array}{c | c}\vA' & \vQ\\\hline \vR & \vS\end{array}\right)$.
\item \label{step:randomization}
For {$i=1,..., m$}:
\subitem Randomly generate $\gamma_i \leftarrow \Fq^{1 \times \alpha}$.
\subitem Set $\va_{i} \leftarrow \va^*_{i}+\gamma_i (\vR||\vS)$
\subitem Set $b_i \leftarrow b^*_i + \gamma_i (b^*_{m+1},..., b^*_{m+k})$
\item Run $D$ on input $\vA, \vb$.
\item Output $D$'s output.
\end{enumerate}
\end{framed}
\caption{A PPT $D'$ that distinguishes LWE using distinguisher for LWE w/ block fixing source}
\label{fig:perfectLWEreduction}
\end{figure}
The distinguisher $D'$ has an advantage when $S$ is of rank $\alpha$.  This occurs with overwhelming probability~(techniques from Cooper~\cite{cooper2000rank}):
\begin{claim}
\label{cl:full rank matrix}
Let $\vS \overset{\$}\leftarrow \Fq^{\alpha \times (\alpha+\beta)}$ be randomly generated.  Then $\Pr[\rank(\vS)=\alpha]\geq  1- \ngl(n)$.\lnote{what is $n$?}
\end{claim}
\begin{proof}
The probability that all $\alpha$ rows are linearly independent is the successive probability that each row is linearly independent of the previous rows, for the $i$th row this probability is $q^{i-1}/q^{\alpha+\beta}$.  By union bound and $\beta = \omega(1)$, the probability that $\vS$ is not full rank is:
\begin{align*}
\Pr[\rank(\vS) < \alpha] &= \sum_{i=1}^{\alpha} \frac{q^{i-1}}{q^{\alpha+\beta}}
\leq \alpha \frac{q^{\alpha}}{q^{\alpha+\beta} }=\ngl(n)\\
\end{align*}
Thus, $\Pr[\rank(\vS) = \alpha] \geq 1-\ngl(n)$.
\end{proof}
\begin{corollary}
$
|\Pr[D(\vA, \vA\vx+\ve) = 1  | \rank(\vS) = \alpha] - \Pr[D(\vA, \vu) =1 | \rank(\vS) = \alpha]|> \epsilon - \ngl(n).
$
\end{corollary}

@@

\subsection{Proof of \lemref{lem:i t poly time}}
\label{sec:proof lem i t poly time}
\begin{proof}
Note that $\decode_t$ will stop if $W$ and $W'$ agree on all the rows selected in Step 2 (it may also stop for other reasons---namely, in step 4; but we do not use this fact to bound the expected running time).
The probability of each selected row having an error is at most $\frac{t}{m - i}$ where $i$ is the number of rows already selected.  That is,
\begin{align*}
\Pr[i_1,..., i_{2n}\text{ have no errors}]&\geq \prod_{i=0}^{2n-1}\left(1 - \frac{t}{m-i}\right)\geq \prod_{i=0}^{2n-1}\left( 1-\frac{d\left(\frac{m}{n}-2\right)\log n}{m-i}\right)\\
&\geq  \prod_{i=0}^{2n-1}\left( 1-\frac{d\log n}{n}\left(\frac{m-2n}{m-i}\right)\right)\geq \prod_{i=0}^{2n-1}\left( 1-\frac{d\log n}{n}\right) \\
&= \left(1-\frac{d\log n}{n}\right)^{2n}  = \left(\left(1-\frac{d\log n}{n}\right)^{\frac{n}{d\log n}}\right)^{2d\log n}\geq \frac{1}{4^{2d\log n}} = \frac{1}{n^{4d}}\,.
\end{align*}
(The second-to-last step holds as long as $n\ge 2d\log n$.) Because at each iteration, we select $2n$ rows independently at random, the expected number of iterations is at most $n^{4d}$; each iteration takes $O(n^3)$ operations in $\Fq$, which gives us the expected running time bound.

The probability that $\decode_t$ outputs $\perp$ is bounded by 
\begin{eqnarray*}
\Pr[\decode_t\rightarrow \perp]& \le & \sum_{j=1}^\infty \Pr[\decode_t\rightarrow \perp \text{ in $j$th iteration of step 4}]\\
&= &\sum_{j=1}^\infty  \Pr[\decode_t \text{ has not stopped after $j-1$ iterations} \wedge \rank(\vA_{i_1},\dots, \vA_{i_{2n}})<n]\\
&\le &\sum_{j=1}^\infty  \Pr[i_1,..., i_{2n}\text{ had errors $j-1$ times} \wedge \rank(\vA_{i_1},\dots, \vA_{i_{2n}})<n]\\
&\le &\sum_{j=1}^\infty  \left(1-\frac{1}{n^{4d}}\right)^{j-1} \cdot e^{-\Omega(n)}\\
& = & n^{4d} e^{-\Omega(n)} = e^{-\Omega(n)}\,.
\end{eqnarray*}
(the second to last line follows from \clref{cl:full rank matrix} \lnote{actually, it doesn't -- \clref{cl:full rank matrix} doesn't have an exponential bound}).

We claim that if the code generated by $\vA$ 
has distance at least $2t+1$, then $\decode_t$ will output $\perp$ or the correct $X'=X$.
Indeed, suppose $X'\neq X$. Since $\vA (X-X')$ has at least $2t+1$ nonzero coordinates by the minimum distance of the code generated by $\vA$, and at most  $t$ of those can be zeroed out by the addition of  $W-W'$, such an $X'$ will not pass Step 6.

The probability that the code generated by $\vA$ has distance lower than $2t+1$ is at most $e^{-\Omega(n)}$ (see \corref{cor:code high distance}), and so is the probability of ouptutting $\perp$, computed above.  This gives the correctness bound for $\decode_t$.
\end{proof}
